# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

```

## Loading and preprocessing the data
```{r acti, results='markup'}
setwd('/Users/gijsbartholomeus/Documents/Datasciencecoursera/Reproducible_data/Week2/')
acti <- read.csv(file = 'activity.csv')
actidate <- acti %>% select(date, steps) %>% na.omit() %>% group_by(date) %>% summarize(daysteps=sum(steps)) 
```

## Creating a histogram with steps per day
```{r}
hist(actidate$daysteps,xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```

## Calculate the mean and median steps per day
```{r}
mean(actidate$daysteps)
median(actidate$daysteps)
```

## plot line with steps per time interval
```{r}
actiinterval <- acti %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(daysteps=mean(steps)) 
ggplot(actiinterval, aes(x=interval, y=daysteps))+ geom_line()
```

## See which day has the maximum daysteps value
```{r}
actiinterval[which(actiinterval$daysteps== max(actiinterval$daysteps)),]
```

## calculate the total sum of missing values
```{r}
sum(is.na(acti))
```

## replace missing values with mean of column
```{r}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
NAmean <- acti%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
```

## create histogram with total steps day with filled in mean
```{r}
FullSummedDataByDay <- aggregate(NAmean$steps, by=list(NAmean$date), sum)
names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```

compare old and new median and mean
```{r}
oldmean <- mean(actidate$daysteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)

oldmedian <- median(actidate$daysteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)

oldmean
newmean
oldmedian
newmedian
```

create a factor variable with weekend or weekday
```{r}
NAmean$WeekendOrWeekday <- ifelse(weekdays(as.Date(NAmean$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "Weekday", "Weekend")
FinalWeek <- NAmean %>% group_by(interval, WeekendOrWeekday) %>% summarise(Mean = mean(steps))
```

create a final blue colored line plot with total steps per interval for weekdays and weekend
```{r}
ggplot(FinalWeek, aes(interval,Mean)) + geom_line(color = "#00AFBB") + facet_grid(WeekendOrWeekday ~ .) + xlab("Interval") + ylab("Mean of Steps") + ggtitle("Comparison of Average Number of Steps in Each Interval")
```
